% SPDX-License-Identifier: CC-BY-4.0
\section{Background and related work}
\label{sec:background}

\subsection{Behavioral host intrusion detection}

Forrest's ``sense of self''~\cite{forrest1996sense} observed that a
process's syscall sequence under normal operation is a stable,
low-entropy fingerprint and that deviations from that fingerprint are
detectable. Subsequent work moved along two axes. \emph{Richer
features}: from raw syscall sequences to file-access profiles,
network endpoints, and child-spawn graphs. \emph{Richer models}: from
fixed-length n-grams to hidden-Markov models, recurrent networks, and
provenance-graph anomaly detectors~\cite{goyal2023mimicry}. The
modeling choices have moved a lot in thirty years. The assumption
that the learning window is adversary-free has not.

\subsection{Mimicry attacks}

A long line of work shows that detectors which inspect only the live
behavioral stream fall to mimicry attacks: an adversary who knows the
detector's signature pads malicious activity with otherwise benign
behavior to evade detection
~\cite{wagner2002mimicry,goyal2023mimicry}. Goyal et al.\ force
misclassification on every provenance-graph host intrusion detector
they test by embedding substructures from legitimate process activity
into the attack subgraph~\cite{goyal2023mimicry}.

Baseline poisoning is related but distinct. Mimicry adversaries shape
what their attack \emph{looks like} at detection time. Poisoning
adversaries shape what the detector considers \emph{normal} at
training time. The two compose: an adversary who can do both has more
leverage than one who can only do one.

\subsection{Robust statistics}

Robust statistics gives several primitives for estimating central
tendency and dispersion in the presence of
outliers~\cite{huber1964robust,rousseeuw1985mvd}. \sediment{} draws
on the trimmed-mean tradition. The classical breakdown point of the
symmetric $\trim$-trimmed mean is $\trim/2$: an adversary controlling
fewer than $\trim N / 2$ samples cannot move the trimmed mean an
arbitrary amount~\cite{huber1964robust}. The piece we add is a
closed-form Gaussian-truncation correction that recovers an unbiased
estimate of the underlying $\sigma$ on clean data while keeping the
trimmed mean's resistance.

\subsection{Concept drift versus adversarial drift}

Production behavioral detectors often conflate two distinct things
under the label ``drift'': natural \emph{concept drift} (the workload
genuinely changed) and \emph{adversarial drift} (an attacker is
manipulating the inputs). \chronosynd{} treats them as orthogonal.
The maintenance-window mechanism in \cref{sec:system} addresses
concept drift by giving operators an explicit way to mark intervals
where baseline shifts are expected. \sediment{} addresses
adversarial drift at the estimator level.
