% SPDX-License-Identifier: CC-BY-4.0
\section{Threat model}
\label{sec:threat-model}

\paragraph{Setting.} A behavioral HIDS observes a per-process feature
stream and fits a baseline from $N$ benign-window observations
$X = \{x_1, \dots, x_N\} \subset \mathbb{R}^d$. After fitting, it
scores live observations against the fitted baseline and raises an
alert when the score exceeds a threshold $\threshold$.

\paragraph{Adversary capability.} The adversary controls a fraction
$\budget \in [0, 1]$ of the learning window. They may inject up to
$|A| = \lfloor \budget N \rfloor$ observations of their choosing,
giving the defender $X' = (X \setminus B) \cup A$ where $B$ is some
discarded subset of the original benign observations.

\paragraph{Adversary goal.} Given a target attack behavior
$\target \in \mathbb{R}^d$ representing the post-learning malicious
action the adversary intends to perform, the adversary picks $A$ to
minimize the score $\score(\target)$ assigned by the baseline fitted
on $X'$. The adversary wins if $\score(\target) < \threshold$.

\paragraph{Observability.} We consider three variants:
\begin{description}
\item[Black-box] The adversary knows there is a baseline-fitting step
  but not the algorithm or hyperparameters. Realized by
  \texttt{inject\_pre\_seed}.
\item[Grey-box] The adversary knows the algorithm family but not the
  exact hyperparameters. Treated as a special case of white-box for
  evaluation.
\item[White-box] The adversary knows the algorithm and all
  hyperparameters, including \trim{}. Realized by
  \texttt{inject\_targeted}, which grid-searches the per-feature
  injection placement that minimizes $\score(\target)$ after a fresh
  fit.
\end{description}

\paragraph{Out of scope.} Kernel-level compromise that lets the
adversary forge process identity (PID, UID, cgroup) or backdoor the
BPF runtime is out of scope. The threat model assumes the kernel-side
identity surface is trustworthy. Post-fit tampering with the
persisted baseline store is a separate concern, addressed in the
runtime by a SHA-256 hash chain over an append-only audit log
(\cref{sec:system}). It is orthogonal to the algorithmic
contribution of this paper.
