% SPDX-License-Identifier: CC-BY-4.0
% Copyright 2026 The ChronosynD Authors, see LICENSE-CC-BY-4.0

\documentclass[10pt,letterpaper]{article}

% core packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}

% figure search path
\graphicspath{{figures/}}

% bibliography
\usepackage[backend=biber,style=numeric,sorting=none]{biblatex}
\addbibresource{references.bib}

% theorem-like environments
\newtheorem{proposition}{Proposition}

% macros
\newcommand{\sediment}{\textsc{Sediment}}
\newcommand{\chronosynd}{\textsc{ChronosynD}}
\newcommand{\trim}{\ensuremath{\tau}} % trim fraction
\newcommand{\budget}{\ensuremath{\beta}} % adversary budget
\newcommand{\target}{\ensuremath{m^{*}}} % target attack behavior
\newcommand{\score}{\ensuremath{s}} % drift score
\newcommand{\threshold}{\ensuremath{\theta}} % detection threshold
\newcommand{\todo}[1]{{\color{red}\textbf{TODO}: #1}}

% title block
\title{\sediment{}: Poisoning-Resistant Behavioral Baselines\\for Host Intrusion Detection}
\author{Vurmusum%
  \thanks{Independent. Email: \texttt{vurmusum@gmail.com}.}}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Behavioral host intrusion detection systems learn a per-process
baseline from a window of observations and score live behavior as
drift against the baseline. Every such system rests on an implicit
assumption that the learning window is free of adversary activity.
In practice attacker dwell time is nonzero by definition, and a
detector that turns on while the adversary is already on the host
fits a baseline shaped by the adversary. We call this \emph{baseline
poisoning}. We formalize the threat model with an adversary budget
$\budget \in [0, 1]$ that bounds the fraction of the window the
adversary controls, propose \sediment{}, a per-feature trimmed-moment
estimator with a closed-form Gaussian-truncation bias correction and
a single trim-fraction knob \trim{} that gives defensible resistance
up to $\budget = \trim{}/2$, and evaluate across nine experiments.
Under uniform poisoning at $\budget = 0.15$ on an isotropic synthetic
workload, the naive mean-and-stddev estimator loses roughly a factor
of 3.6 in target-score discriminability while \sediment{} at
$\trim{} = 0.30$ holds. Under a grid-search white-box adversary with
full knowledge of \trim{}, the optimal placement reduces to the
at-target injection, which gives evidence that white-box knowledge
does not yield a stronger attack against symmetric-trim defenders.
On a real captured trace of Linux \texttt{bash} activity, \sediment{}
detects an attack-payload run with a 126$\times$ median score
separation over clean windows and zero false positives during the
attack capture period. A reference Python implementation and a
production Rust runtime stay bit-equivalent through
cross-implementation parity tests on every push.
\end{abstract}

\input{sections/intro}
\input{sections/background}
\input{sections/threat_model}
\input{sections/system}
\input{sections/sediment}
\input{sections/evaluation}
\input{sections/discussion}
\input{sections/limitations}
\input{sections/conclusion}

\printbibliography

\end{document}
